Forecasting

### 2022-10-18 manuscript/forecasting_ed.qmd

---
title: Forecasting for emergency medcine
author:
  - name: Author1
    email: emaail@example.com
    affiliation: University1
    footnote: 1
  - name: Author2
    email: email2@example.com
    affiliation: University2
    footnote: 2
address:
  - code: University1
    address: adress1
  - code: University2
    address: adress2
footnote:
  - code: 1
    text: "Corresponding Author"

abstract: |
   The Objective of this paper
keywords:  ""
journal: "Which journal?"
date: "`r Sys.Date()`"
geometry: "top=25mm, left=30mm, right=30mm, bottom=25mm,headsep=10mm, footskip=12mm"
#linenumbers: true
numbersections: true
#output: rticles::elsevier_article
header-includes:  
 \usepackage{adjustbox, float,lscape} #use the 'float' package
output: 
  bookdown::pdf_book:
    base_format: rticles::elsevier_article
    extra_dependencies: ["flafter"]
    keep_tex: true
bibliography: mybibfile.bib
biblio-style: authoryear
#csl: elsevier-without-titles.csl
#biblio-style: abbrevnat
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE, cache=TRUE, message = FALSE, warning = FALSE)
# Make sure you have the latest versions of rmarkdown and rticle
library(tidyverse)
library(lubridate)
library(patchwork)
library(tsibble)
library(kableExtra)
library(fable)
library(ggthemes)
library(readxl)
```

@@

@@

@@

@@

Table \@ref(tab:summarylit) summarise studies on forecasting in emergency and urgent care.

```{r summarylit}
li <- readxl::read_xlsx("../data/literature.xlsx")
knitr::kable(li, booktabs = T, linesep = "",caption = "Summary of studies in hourly emergency care forecasting") %>% kableExtra::kable_styling(latex_options = c("scale_down","hold_position")) %>% column_spec(5,"15em")
```

>>>>

---
title: Forecasting
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Annals of Emergency Medicine
      formatting: preprint
      model: 3p
      cite-style: authoryear
  elsevier-html: 
    toc: true
author:
  - name: Bahman Rostami-Tabar
    email: rostami-tabarb@cardiff.ac.uk
    affiliations: 
        - id: carbs
          name: Cardiff University
          department: Cardiff Business School
          country: United Kingdom
          postal-code: CF10 3EU
    attributes:
        corresponding: true
    note: This is the first author footnote.
  - name: Rob J. Hyndman
    email: Rob.Hyndman@monash.edu
    affiliations:
        - id: monash
          name: Monash University
          department: Department of Econometrics and Business Statistics
          country: Australia
          postal-code: VIC 3800
abstract: |
  This is the abstract. An accurate demand forecasting is crucial in Emergency medicine to depict various courses of action that can result in massive savings in terms of patient lives. Inability to match the staff with the demand results in an overcrowding care system which is a serious problem causing challenging situations on patient flow. Also, it is related with increasing length of stay, low patient satisfaction, increasing health care costs, inaccuracy in electronic medical record, and reported waiting times.
keywords: [Forecasting, Ambulance, Hierarchical structure]
bibliography: mybibfile.bib
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE, cache=TRUE, message = FALSE, warning = FALSE)
# Make sure you have the latest versions of rmarkdown and rticle
library(tidyverse)
library(lubridate)
library(patchwork)
library(tsibble)
library(kableExtra)
library(fable)
library(ggthemes)
library(readxl)
library(here)
```

@@

 The ability to accurately forecast the demand of future health services is critical for planning decisions aimed at improving the quality of service delivery in the health care. Forecasting has been used in various fields of healthcare to inform planning decisions at the strategic, tactical and operational levels. There is at least one or multiple reasons why we need to produce a forecast. Most often reasons determine clearly what to forecast and which level of forecast granularities are required. There are typically some planning process at strategic(long-term) level such as dimensioning resource capacities and budgeting; tactical(medium-term) level such as temporary capacity expansions, staff-shift scheduling, and inventory control; and operational(short-term) level such as planning rosters for staff and allocate resources and medical item distribution. Each planning level will typically have different forecasting needs in terms of granularity and horizon (how far into the future do we forecast). A natural question that arise from these examples is how to generate forecasts to support planning decisions at multiple granularities and horizons. In order to answer this question, we first need to understand some fundamental aspects of forecasting in healthcare:

@@

Implementing and sustaining improvements in hospital-wide flow requires alignment, cooperation, and coordination between hospital units and departments. Without effective executive oversight and collaboration, teams operate in isolation from one another and the aggregated impact of their efforts is limited. In many cases, this isolation leads to duplicative work, rework, or work that runs counter to overall goals to improve hospital-wide patient flow

@@

<!-- An accurate demand forecast by hour of the day enables planners to match staff to meet anticipated patients, reconfigure units and redeploy staff. This will have many advantages for both patients, staff and the quality of provided services. Hourly forecasts are required to inform the short-term operational planning for the current and the upcoming shifts of the day. This involves the short-term decision making related to the execution of the delivery process in ED. The combination of an hourly forecast demand, current staff being occupied, resource availability and waiting times at ED, provide information on the state of the unscheduled care system across the service. Having this full picture enables the delivery managers to focus on the areas that require intervention to enable the most effective delivery of the service to the patients. Moreover, there are often unplanned events, such as walk-in patients, extended consultation times, during a day or a week. Staff capacities may be adjusted based on the predicted demand fluctuations by using part-time, on-call nurses, staff overtime and voluntary absenteeism. That may also require to adjust the scheduling. -->

Despite a large number of studies dedicated to forecasting in healthcare[ref], this ques-
tion unanswered in healthcare forecasting. Almost all studies assume that the level of
data granularity matches with the level of forecast requirement, e.g. daily series is used
to generate daily forecast. Moreover, in the presence of the high data granularity, the
recommendation is to first aggregate the data into a requested forecast granularity level
and then generate the forecast(Goodwin, 2018). However, there is no empirical evidence
to support this claim

@@

Table @tbl-lit summarise studies on forecasting in emergency and urgent care.

```{r}
#| label: tbl-lit
#| tbl-cap: "summary of literature review"
lit <- readxl::read_xlsx(here("data/literature.xlsx"))
 lit %>% kable() %>%  kableExtra::kable_styling(latex_options = c("scale_down","hold_position")) %>% column_spec(5,"15em")
```
